Mechanized Reasoning about a Capability Machine

### 2019-10-18 prisc-2020/paper.tex
added a final conclusion, changed from RO to RX, added a line about the size of the project in the introduction, only missing is to count the number of lines of code

In summary, we present Iris formalizations of 
\begin{itemize}
  \item a program logic for reasoning about capability machine programs.
  \item the logical relation from \cite{skorstengaardESOP18}, 
    which captures capability safety and which can be used to reason about examples
    that rely on WBCF and LSE.
\end{itemize}
All of the technical development is mechanized in Coq using the Coq implementation of Iris.

\todo{Alix: I think we need a short discussion (not specifically here) about the fact that we don't really have a theorem that says something like "being in the logical relation implies LSE and WBCF" as that's not really well defined, but we can prove things about specific programs that rely on it}
\todo{Lars: I just tried to reduce the claim above}

@@

Next we use the Iris framework to \emph{prove} a Hoare triple for each instruction.
As in \cite{Jensen:2013}, we define a points-to predicate for registers, denoted $r \mapsto_r w$.
Since a capability machine replaces pointers with capabilities, we replace the conventional points-to predicate of separation logic for pointers with a points-to predicate with a permission attached to it, denoted
$a \mapsto_a[p]~w$. 
% \[
%    \hspace{1cm} a \mapsto_a[p]~w.
% \]
This predicate states that address $a$ points to word $w$ with permission $p$.
The permission restricts how the memory may be updated at address $a$.
For instance, if $a \mapsto_a[\textsc{\tiny RO}]~w$, then the $\textsc{\tiny RO}$ (ReadOnly)
permission only gives us permission to read from address $a$, but not to write to it.

@@

This notion of World is different from the logical relation given by \cite{skorstengaardESOP18}, where the World is a collection of invariants describing the behaviour of all memory, not just local state.
We can use a simpler more abstract notion of world because XXX
\todo{Lars: say why succintly, is it just because we use capability points-to predicates ?
  But that could also have been done before\ldots}

\todo{Lars: the following is too hard to read and understand -- I suggest to omit it and just add a succinct description above}
% \todo{begin}
% Skorstensgaard et.
% al.
% use the World to define validity of capabilities, by relating the invariant of the region in question to certain standard islands for capabilities.

@@

XXX Add that we prove FTLR and that the LR can be used to prove examples that rely
on WBCF and LSE.

>>>>

In summary, we present Iris formalizations of 
\begin{itemize}
  \item a program logic for reasoning about capability machine programs.
  \item the logical relation from \cite{skorstengaardESOP18}, 
    which captures capability safety and which can be used to reason about examples
    that rely on WBCF and LSE.
\end{itemize}
Most of the technical development is mechanized in Coq using the Coq implementation of Iris (the remaining work consists of a few missing details). Currently the mechanization consists of XXX lines of code. Its substantial size can be slightly improved by some clever use of tactics, however such a mechanization will always be large, due to the low level nature of a machine: we want to consider every possible case of an instruction, both the fails and the successes.

%\todo{Alix: I think we need a short discussion (not specifically here) about the fact that we don't really have a theorem that says something like "being in the logical relation implies LSE and WBCF" as that's not really well defined, but we can prove things about specific programs that rely on it}
%\todo{Lars: I just tried to reduce the claim above}

@@

Next we use the Iris framework to \emph{prove} a Hoare triple for each instruction.
As in \cite{Jensen:2013}, we define a points-to predicate for registers, denoted $r \mapsto_r w$.
Since a capability machine replaces pointers with capabilities, we replace the conventional points-to predicate of separation logic for pointers with a points-to predicate with a permission attached to it, denoted
$a \mapsto_a[p]~w$. 
% \[
%    \hspace{1cm} a \mapsto_a[p]~w.
% \]
This predicate states that address $a$ points to word $w$ with permission $p$.
The permission restricts how the memory may be updated at address $a$.
For instance, if $a \mapsto_a[\textsc{\tiny RX}]~w$, then the $\textsc{\tiny RX}$ (ReadExecute)
permission gives us permission to read from address $a$, to execute w, but it does not to permit us to write to address $a$.

@@

This notion of World is different from the logical relation given by \cite{skorstengaardESOP18}, where the World is a collection of invariants describing the behaviour of all memory, not just local state.
We can use a simpler more abstract notion of world because the Iris model takes care of the world circularity problem. The necessary step indexing is abstracted away by the Iris saved predicates, which we use to define the validity of the regions a capability has authority over. 
%\todo{Lars: say why succintly, is it just because we use capability points-to predicates ?
%  But that could also have been done before\ldots}

%\todo{Lars: the following is too hard to read and understand -- I suggest to omit it and just add a succinct description above}
% \todo{begin}
% Skorstensgaard et.
% al.
% use the World to define validity of capabilities, by relating the invariant of the region in question to certain standard islands for capabilities.

@@

We prove the Fundamental Theorem of Logical Relations, in which we show that if we can read a region, we can execute it. We then use the fundamental theorem to prove examples with calls to unknown adversary code, which rely on local state encapsulation and well bracketed control flow.
